figures_manuscript

### 2025-01-04 pub_figures/manuscript-figures.Rmd
NMR plot v1

---
title: "figures_manuscript"
output:
  github_document:
    html_preview: false
    toc: true
    toc_depth: 2
knit: (function(input_file, encoding) {
    rmarkdown::render(input_file,output_file= 'README.md')
        })
date: "2024-10-17"
editor_options: 
  chunk_output_type: console
---

@@

# Fig 3: Volcanos
## a) Genotype (1)
```{r volcanos1, fig.height=6, fig.width=10}

@@

ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Volcanos.pdf",sep=""), plot = volcanos_all,
width = 10,
height = 6)

@@

## b): light effect (2)
```{r volcanos2, fig.height=6, fig.width=10}

@@

# ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Volcanos.pdf",sep=""), plot = volcanos_all,
# width = 10,
# height = 6)

@@

DEGs_genes["dark"]
blue_red_genes <- unlist(DEGs_genes[c("blue","red")]) %>% unique()

non_sig <- data.frame(x = c(-9, 0, 9), y = c(20, 0.5, 20))

DEGs_comb$sig

gg_all <- ggplot(DEGs_comb, aes(x=l2FC.blue, y=l2FC.red, label=symbol)) + # color=group, fill=group 
  geom_point(shape=21) +
  geom_hline(yintercept = c(-1,1), linewidth = 0.1) + 
  geom_vline(xintercept = c(-1,1), linewidth = 0.1) +
  coord_cartesian(xlim=c(-10,10),ylim = c(-10,10)) +
  geom_text_repel(size=2, max.overlaps = 20) +
  theme_bw() +
  removeGrid(x=T, y=T)
gg_all

gg_br <- ggplot(DEGs_comb[blue_red_genes,], aes(x=l2FC.blue, y=l2FC.red, label=symbol)) + # color=group, fill=group 
  geom_point(shape=21) +
  geom_text_repel(size=3) +
  coord_cartesian(xlim=c(-10,10),ylim = c(-10,10)) +
  theme_bw() +
  removeGrid(x=T, y=T)
gg_br

red_specific <- DEGs_comb %>% subset(l2FC.blue < 1 & l2FC.blue > -1 & (l2FC.red > 0.5 | l2FC.red < -0.5) & padj.red < 0.05) %>% arrange(l2FC.red)

goi <- "Cre03.g192050"

# single gene
d <- plotCounts(dds, gene=goi, intgroup=c("condition","experiment","genotype","treatment"), main=s,returnData=TRUE)
colnames(d)[1] <- "counts"

gcounts <- ggplot(d, aes(x = treatment, y = counts, fill=condition, color=condition)) +
    geom_boxplot(color="black") +
    geom_point(shape=21,color="black",aes(fill=condition),position=position_dodge(width=0.75), alpha=1) +
  # geom_text_repel(aes(label=clientName)) +
    scale_fill_manual(values=group.colors) +
    scale_color_manual(values=group.colors) +
    scale_y_continuous(trans = "log2") +
    theme_bw() +
    removeGrid(x=T, y=T) +
    geom_vline(xintercept=seq(1,length(levels(all_counts$treatment))-1,1)+.5,color="grey") +
    labs(title = paste(mcols(dds)[goi,]$symbol," in ",colData(dds)$experiment[1],sep = "")) +
    theme(axis.text.x = element_text(size = 10, face = "bold"),
    axis.text.y = element_text(color = "grey20", size = 8, angle = 0, hjust = 1, vjust = 0, face = "plain"),
    axis.title.x = element_text(color = "white", size = 12, angle = 0, hjust = .5, vjust = 0, face = "plain"),
    axis.title.y = element_text(color = "grey20", size = 12, angle = 90, hjust = .5, vjust = .5, face = "bold"))
gcounts

>>>>

---
title: "figures_manuscript"
output:
  github_document:
    html_preview: false
    toc: true
    toc_depth: 2
knit: (function(input_file, encoding) {
    rmarkdown::render(input_file,output_file= 'README.md')
        })
date: "2024-10-17"
editor_options: 
  chunk_output_type: inline
---

@@

# Fig 3: Volcanos
## a) Δgenotype per condition (1)
```{r volcanos1, fig.height=6, fig.width=10}

@@

ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Volcanos1.pdf",sep=""), plot = volcanos_all,
width = 10,
height = 6)

@@

## b): Δlight = light effect (2)
```{r volcanos2, fig.height=10, fig.width=10}

@@

ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Volcanos2.pdf",sep=""), plot = volcanos_2,
width = 10,
height = 12)

## c): ΔgenotypeΔlight = difference in light effect (interaction term) (3)
```{r volcanos3, fig.height=6, fig.width=10}

res <- res_ashr_list$pcry$pcry_BLvD.vs.WT_BLvD
res_n <- res_list$pcry$pcry_BLvD.vs.WT_BLvD

volcano3_blue <- EnhancedVolcano(res,
    lab = mcols(dds_list[["pcry"]])[["geneSymbol"]],
    x = 'log2FoldChange',
    y = 'padj',
    col=c("grey","grey","grey",group.colors[3]),
    title = "pcry_BLvD.vs.WT_BLvD",
    titleLabSize = 12,
    subtitle = paste0("upregulated: ",up,", downregulated: ",down,"\n(total: ",total,")"),
#    subtitle = {},
    subtitleLabSize = 10,
    caption = NULL,
    xlim = c(-7,7),
    ylim = c(0,50),
    pCutoff = 0.05,
    FCcutoff = 1,
    maxoverlapsConnectors = 50,
    drawConnectors = TRUE,
    widthConnectors = 0.5,
    colConnectors = "grey70",
    legendLabels=c('ns','ns','ns',
      'padj < 0.05 & Log2FC > 1'),
    labSize = 4,
    axisLabSize = 12,
    legendLabSize = 12,
    legendIconSize = 3,
    gridlines.major = FALSE,
    gridlines.minor = FALSE,
    pointSize = 3
)

# volcano_red
res_ashr_list %>% names()
res_ashr_list[[2]] %>% names()

res <- res_ashr_list$pcry$pcry_RvD.vs.WT_RvD
res_n <- res_list$pcry$pcry_RvD.vs.WT_RvD

res[res$log2FoldChange > l2FCmax,]$log2FoldChange <- l2FCmax
res[res$log2FoldChange < -l2FCmax,]$log2FoldChange <- -l2FCmax
x <- res$padj
res[(x < pmax) & (!is.na(x)),]$padj <- pmax

volcano3_red <- EnhancedVolcano(res,
    lab = mcols(dds_list[["pcry"]])[["geneSymbol"]],
    x = 'log2FoldChange',
    y = 'padj',
    col=c("grey","grey","grey",group.colors[5]),
    title = "pcry_RvD.vs.WT_RvD",
    titleLabSize = 12,
    subtitle = paste0("upregulated: ",up,", downregulated: ",down,"\n(total: ",total,")"),
#    subtitle = {},
    subtitleLabSize = 10,
    caption = NULL,
    xlim = c(-7,7),
    ylim = c(0,50),
    pCutoff = 0.05,
    FCcutoff = 1,
    maxoverlapsConnectors = 50,
    drawConnectors = TRUE,
    widthConnectors = 0.5,
    colConnectors = "grey70",
    legendLabels=c('ns','ns','ns',
      'padj < 0.05 & Log2FC > 1'),
    labSize = 4,
    axisLabSize = 12,
    legendLabSize = 12,
    legendIconSize = 4,
    gridlines.major = FALSE,
    gridlines.minor = FALSE,
    pointSize = 3
)

volcanos_3 <- (volcano3_blue + volcano3_red) +
  plot_layout(guides = "collect", axes="collect", axis_titles="collect") & 
  theme(legend.position = 'none', axis.title=element_text(size=12))
volcanos_3

ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Volcanos3.pdf",sep=""), plot = volcanos_3,
width = 10,
height = 6)

colnames(anno)
anno[deg_list$pcry$deg_pcry_RvD.vs.WT_RvD,]
anno[deg_list$pcry$deg_pcry_RvD.vs.WT_RvD,c("geneSymbol","previousIdentifiers","Description","previousIdentifiers","Description","Comments","TMHMM_transmembrane","TargetP","Predalgo","Flagellar_Proteome")] %>% kable()

@@

goi_tf

res_core <- list(WT_blue = res_list$pcry$WT_BL.vs.D,
             WT_red = res_list$pcry$WT_R.vs.D,
             pcry_blue = res_list$pcry$pcry_BL.vs.D,
             pcry_red = res_list$pcry$pcry_R.vs.D,
             dd_blue = res_list$pcry$pcry_BLvD.vs.WT_BLvD,
             dd_red = res_list$pcry$pcry_RvD.vs.WT_RvD) %>% lapply(data.frame)

# plot(res_list$pcry$pcry_BL.vs.D$log2FoldChange~res_list$pcry$pcry_R.vs.D$log2FoldChange)
#res_list$pcry$WT_BL.vs.D$symbol == anno[res_list$pcry$WT_BL.vs.D %>% rownames(),"geneSymbol"]

res_comb <- bind_cols(symbol = anno[res_list$pcry$WT_BL.vs.D %>% rownames(),"geneSymbol"],
                       res_core[[1]][,c(1,2,6)],
                       res_core[[2]][,c(2,6)],
                       res_core[[3]][,c(2,6)],
                       res_core[[4]][,c(2,6)],
                       res_core[[5]][,c(2,6)],
                       res_core[[6]][,c(2,6)])

colnames(res_comb)[c(3,4)] <- c("l2FC.WT_blue","padj.WT_blue")
colnames(res_comb)[c(5,6)] <- c("l2FC.WT_red","padj.WT_red")
colnames(res_comb)[c(7,8)] <- c("l2FC.pcry_blue","padj.pcry_blue")
colnames(res_comb)[c(9,10)] <- c("l2FC.pcry_red","padj.pcry_red")
colnames(res_comb)[c(11,12)] <- c("l2FC.dd_blue","padj.dd_blue")
colnames(res_comb)[c(13,14)] <- c("l2FC.dd_red","padj.dd_red")

res_comb$symbol2 <- anno[res_comb %>% rownames(),"geneSymbol"]
  blue_red_genes <- c(deg_list$pcry$deg_pcry_BLvD.vs.WT_BLvD,
                    deg_list$pcry$deg_pcry_RvD.vs.WT_RvD) %>%
  unique()

non_sig <- data.frame(x = c(-20, 0, 20), y = c(20, 0.5, 20))

df <- res_comb[anno_tf$gene_id,]
df <- res_comb[anno_phot$gene_id,]

gg_phots <- ggplot(df, aes(x=l2FC.WT_blue, y=l2FC.WT_red, label=symbol)) + # color=group, fill=group 
    geom_segment(aes(x = l2FC.WT_blue, y= l2FC.WT_red, xend = l2FC.pcry_blue,yend = l2FC.pcry_red, color = l2FC.dd_red),
        arrow = arrow(length = unit(0.01,units = "npc"))) +
  scale_color_gradient2(mid="grey") +
  geom_point(shape=21, fill="grey", col="grey40") +
  geom_point(aes(x=l2FC.pcry_blue, y=l2FC.pcry_red),shape=21, fill="green3", col="grey40") +
  geom_hline(yintercept = c(-1,1), linewidth = 0.1) + 
  geom_vline(xintercept = c(-1,1), linewidth = 0.1) +
  geom_abline(slope=c(1), intercept = 0, linewidth = 0.1) +
  # coord_cartesian(xlim=c(-5,5),ylim = c(-5,5)) +
  coord_fixed() +
  geom_text_repel(size=4, max.overlaps = 20) +
  theme_bw() +
  removeGrid(x=T, y=T)
gg_phots

# ROCs
df <- res_comb[anno_tf$gene_id,]
gg_rocs <- ggplot(df, aes(x=l2FC.WT_blue, y=l2FC.WT_red, label=symbol)) + # color=group, fill=group 
    geom_segment(aes(x = l2FC.WT_blue, y= l2FC.WT_red, xend = l2FC.pcry_blue,yend = l2FC.pcry_red, color = l2FC.dd_red),
        arrow = arrow(length = unit(0.01,units = "npc"))) +
  scale_color_gradient2(mid="grey") +
  geom_point(shape=21, fill="grey", col="grey40") +
  geom_point(aes(x=l2FC.pcry_blue, y=l2FC.pcry_red),shape=21, fill="green3", col="grey40") +
  geom_hline(yintercept = c(-1,1), linewidth = 0.1) + 
  geom_vline(xintercept = c(-1,1), linewidth = 0.1) +
  geom_abline(slope=c(1), intercept = 0, linewidth = 0.1) +
  # coord_cartesian(xlim=c(-5,5),ylim = c(-5,5)) +
  coord_fixed() +
  geom_text_repel(size=4, max.overlaps = 20) +
  theme_bw() +
  removeGrid(x=T, y=T)
gg_rocs

# TOP
df <- res_comb[blue_red_genes,]

gg_top <- ggplot(df, aes(x=l2FC.WT_blue, y=l2FC.WT_red, label=symbol)) + # color=group, fill=group 
    geom_segment(aes(x = l2FC.WT_blue, y= l2FC.WT_red, xend = l2FC.pcry_blue,yend = l2FC.pcry_red, color = l2FC.dd_red),
        arrow = arrow(length = unit(0.01,units = "npc"))) +
  scale_color_gradient2(mid="grey") +
  geom_point(shape=21, fill="grey", col="grey40") +
  geom_point(aes(x=l2FC.pcry_blue, y=l2FC.pcry_red),shape=21, fill="green3", col="grey40") +
  geom_hline(yintercept = c(-1,1), linewidth = 0.1) + 
  geom_vline(xintercept = c(-1,1), linewidth = 0.1) +
  geom_abline(slope=c(1), intercept = 0, linewidth = 0.1) +
  coord_cartesian(xlim=c(-5,5),ylim = c(-5,5)) +
  # coord_fixed() +
  geom_text_repel(size=4, max.overlaps = 20) +
  theme_bw() +
  removeGrid(x=T, y=T)
gg_top
